Dynamical Neural Network Models\\for Sequence Learning

### 2019-05-07 paper/main.tex

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage[backend=biber,sorting=none]{biblatex}
%\usepackage{geometry}
\addbibresource{mybib.tex}

@@

\section*{Background}
There have been some recent studies comparing various models for time-series prediction tasks \cite{cui2016continuous} on discretized data. In this study, we sill replicate their results to establish a baseline and propose unsupervised spiking neural architectures for performing this task. Spiking neurons \cite{gerstner2002spiking} are biologically-inspired models of neurons. Spiking neural networks (SNNs) \cite{maass1997networks} are organized connections between these spiking neurons which have more energy efficient \cite{cruz2012energy} implementations on neuromorphic hardware.

@@

\begin{figure}[!h]
    \centering
    \includegraphics[width=\linewidth]{../notebooks/matrix-distances.png}
    \caption{Heatmap of squared errors between distributed random encodings for the sequence characters, with $-1$ as the separating bit for ease of visualization.}
    \label{fig:encoding-distance}
\end{figure}

@@

To present each of these discrete categories to the model, we have to encode the categories to a vector representations. One possibility is one-hot encoding where the vector is the same size the number of unique characters in the sequence, but this representation scales poorly when the number of unique characters increases. For this reason, the characters are encoded as random vectors with real values in the interval $\left[-1, 1\right]$, similar to distributed representations used in natural language learning. \cite{mikolov2013distributed} The euclidean distances between pairs of unique sequence characters using a $25$ length random ending is shown in Figure \ref{fig:encoding-distance}. The decoder for converting vectors back to character uses a nearest neighbor approach, therefore the precision in the output vector from the model is important as the number of noise characters increases. Another possibility to consider for the input encoding would be the sparse distributed categorical encoder utilized by Cui et al. (2016) for their hierarchical temporal memory (HTM) model.

@@

Long short-term memory \cite{hochreiter1997long} networks are able to achieve state of the art results on various sequence learning tasks, so they are a good starting point to establish baseline performance on this task. We use the same LSTM architecture proposed by Cui et al. (2016) with $25$ input neurons connected to a hidden layer of $20$ LSTM cells and $25$ output neurons. The network is trained using truncated backpropagation through time (TBPTT) \cite{mozer1995focused} on the last $100$ seen elements. The output is classified using the nearest neighbor decoder.

\subsubsection*{LSTM Behavior}
There are many different variants of LSTM implementations. In this work, the LSTM cell has a forget gate, but no peephole connections. The equations defining the forward pass of the hidden layer in the LSTM model are:

@@

In the above equations, $\sigma$ is the sigmoid activation function, $i_t$ is the input gate, $f_t$ is the forget gate, $g_t$ is the cell gate, $o_t$ is the output gate and the $c_t$ and $h_t$ are the cell and hidden states respectively. The $\ast$ operator is the elementwise product operator. The biases $\left(b_i, b_f, b_g, b_o\right)$ and weights $\left(W_{ii}, W_{hi}, W_{if}, W_{hf}, W_{ig}, W_{hg},W_{io}, W_{ho}\right)$ are initialized uniformly from $\left[-\frac{1}{k}, \frac{1}{k}\right]$, where $k$ is the number of hidden units. These terms are then learned through truncated backpropagation through time (TBPTT) which involves keeping a short history of elements and clipping the gradient updates to the beginning of the history. \cite{sutskever2013training}

\begin{figure}[!h]
    \centering
    \includegraphics[width=\linewidth]{../diagrams/lstm.png}
    \caption{The network architecture for the LSTM-online model. The LSTM cells maintain an internal hidden state.}
    \label{fig:lstm-online-model}
\end{figure}

@@

Time-delay neural networks (TDNN) are feed-forward neural networks trained using a lag on a window of previous data. \cite{waibel1995phoneme} For this task, the TDNN is trained on the last $1000$ characters using a lag of $10$ using backpropagation. \cite{rojas1996backpropagation} The model has $250$ input neurons that are fully connected to a hidden layer of $200$ neurons. The hidden layers utilize a ReLu non-linearity and are fully connected to $25$ output neurons. This model proposed by Cui et al. have a sigmoid non-linearity in the hidden layer. We replaced the non-linearity to allow for conversion into a spiking neural network. As with the LSTM model, the output is classified using the nearest neighbor decoder.

\begin{figure}[!h]
    \centering
    \includegraphics[width=\linewidth]{../diagrams/tdnn.png}
    \caption{The network architecture for the TDNN/TDSNN model.}
    \label{fig:lstm-online-model}
\end{figure}

@@

A common approach to training spiking neural networks \cite{diehl2015fast} is through the transfer of weights learned through backpropagation on an identical non-spiking neural network and then scaling the weights. Using this approach, we can convert the TDNN model trained with backpropagtion into an equivalent spiking network using the approach developed by Diehl et al. (2016) This method doesn't perform well when the output values are negative. To facilitate these negative spikes, the output layer was doubled in size and the negative weight matrix of the original connections was concatentated. The same was done for the bias. The positive sum of spikes was subracted from the negative sum of spikes and then divided by the runtime to obtain an approximation of the original network's value.
The drawback of using this method when we need to approximate real-values is the loss of precision based on the duration of the input. For this task, we show the input for $500$ simulated milliseconds. As with the other models, the output is classified using the nearest neighbor decoder.

@@

While conversion is one way to train a neural network, it is not energy efficient to perform backpropagation at every time step and then convert. We propose a spiking neural network architecture, with $10$ inputs, each fully connected to $1$ of $10$ columns of $100$ spiking neurons. Each of the columns has a Hebbian connection to each of the other columns. To map the spikes from this network, we train a  K-Nearest Neighbor model using previous spikes and values. \cite{beliaev2007time} The model diagram is shown in Figure \ref{fig:csnn-network-diagram}.

\begin{figure}[!h]
    \centering
    \includegraphics[width=\linewidth]{../diagrams/csnn.png}
    \caption{The network architecture for the CSNN model.}
    \label{fig:csnn-network-diagram}
\end{figure}

@@

@@

%\subsection*{Perturbation}

%Another important property of each of these models is the robustness to temporal noise. We introduce an $\alpha$ parameters which represents the probability that a character is swapped for another character.

@@

%\section*{Conclusion}

>>>>

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage[backend=biber,sorting=none]{biblatex}
\usepackage{geometry}
\addbibresource{mybib.tex}

@@

\section*{Background}
There have been some recent studies comparing various models for time-series prediction tasks \cite{cui2016continuous} on discretized data. In this study, we will replicate their results to establish a baseline and propose unsupervised spiking neural architectures for performing this task. Spiking neurons \cite{gerstner2002spiking} are biologically-inspired models of neurons. Spiking neural networks (SNNs) \cite{maass1997networks} are organized connections between these spiking neurons which have more energy efficient \cite{cruz2012energy} implementations on neuromorphic hardware.

@@

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.7\linewidth]{../notebooks/matrix-distances.png}
    \caption{Heatmap of squared errors between distributed random encodings for the sequence characters, with $-1$ as the separating bit for ease of visualization.}
    \label{fig:encoding-distance}
\end{figure}

@@

To present each of these discrete categories to the model, we have to encode the categories to a vector representations. One possibility is to one-hot encode each character, but this representation scales poorly when the number of unique characters increases. For this reason, the characters are encoded as random vectors with real values in the interval $\left[-1, 1\right]$, similar to distributed representations used in natural language learning. \cite{mikolov2013distributed} The euclidean distances between pairs of unique sequence characters using a $25$ length random ending is shown in Figure \ref{fig:encoding-distance}. The decoder for converting vectors back to character uses a nearest neighbor approach, therefore the precision in the output vector from the model is important as the number of noise characters increases.

@@

Long short-term memory \cite{hochreiter1997long} networks are able to achieve state of the art results on various sequence learning tasks, so they are a good starting point to establish baseline performance on this task. We use the same LSTM architecture proposed by Cui et al. (2016) with $25$ input neurons connected to a hidden layer of $20$ LSTM cells and $25$ output neurons. The network is trained using truncated backpropagation through time (TBPTT) \cite{mozer1995focused, sutskever2013training} on the last $100$ seen elements. The output is classified using the nearest neighbor decoder.

\subsubsection*{Long short-term memory (LSTM) Behavior}
There are many different variants of LSTM implementations. In this work, the LSTM cell has a forget gate, but no peephole connections. The equations defining the forward pass of the hidden layer in the LSTM model are:

@@

In the above equations, $\sigma$ is the sigmoid activation function, $i_t$ is the input gate, $f_t$ is the forget gate, $g_t$ is the cell gate, $o_t$ is the output gate and the $c_t$ and $h_t$ are the cell and hidden states respectively. The $\ast$ operator is the elementwise product operator. The biases $\left(b_i, b_f, b_g, b_o\right)$ and weights $\left(W_{ii}, W_{hi}, W_{if}, W_{hf}, W_{ig}, W_{hg},W_{io}, W_{ho}\right)$ are initialized uniformly from $\left[-\frac{1}{\sqrt{k}}, \frac{1}{\sqrt{k}}\right]$, where $k$ is the number of hidden units. \cite{jozefowicz2015empirical}

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.9\linewidth]{../diagrams/lstm.png}
    \caption{The network architecture for the LSTM-online model. The LSTM cells maintain an internal hidden state.}
    \label{fig:lstm-online-model}
\end{figure}

@@

Time-delay neural networks (TDNN) are feed-forward neural networks trained using a lag on a window of previous data. \cite{waibel1995phoneme} For this task, the TDNN is trained on the last $3000$ samples every $1000$ elements with a lag of $10$. \cite{rojas1996backpropagation} The model has $250$ input neurons that are fully connected to a hidden layer of $200$ neurons. The hidden layers utilize the ReLU non-linearity and are fully connected to $25$ output neurons. The model proposed by Cui et al. has a sigmoid non-linearity in the hidden layer. We replaced the non-linearity to allow for conversion into a spiking neural network. As with the LSTM model, the output is classified using the nearest neighbor decoder.

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.9\linewidth]{../diagrams/tdnn.png}
    \caption{The network architecture for the TDNN/TDSNN model.}
    \label{fig:lstm-online-model}
\end{figure}

@@

A common approach to training spiking neural networks is through the transfer of weights learned through backpropagation on an identical non-spiking neural network and then scaling the weights. \cite{diehl2015fast} Using this approach, we can convert the TDNN model trained with backpropagtion into an equivalent spiking network using the approach developed by Diehl et al. (2016)

This method works reasonably well when the output values of the network are positive, however the output has values that are negative. To facilitate these negative spikes, the output layer was doubled in size and the negative weight matrix of the original connections was concatentated. The same was done for the bias. The positive sum of spikes was subtracted from the negative sum of spikes and then divided by the runtime to obtain an approximation of the original network's output.

The drawback of using this method when we need to approximate real-values is the loss of precision based on the duration of the input. For this task, we show the input for $500$ simulated milliseconds. As with the other models, the output is classified using the nearest neighbor decoder.

@@

While conversion is one way to train a neural network, it is not energy efficient to perform backpropagation and then convert. We propose a spiking neural network architecture, with $10$ inputs, each fully connected to $1$ of $10$ columns of $100$ spiking neurons with randomly initialized weights. Each of the columns has a Hebbian connection to each of the other columns initialized with zero weights. To map the spikes from this network, we train a K-Nearest Neighbor model using previous spikes and values. \cite{beliaev2007time} The model diagram is shown in Figure \ref{fig:csnn-network-diagram}.

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.9\linewidth]{../diagrams/csnn.png}
    \caption{The network architecture for the CSNN model.}
    \label{fig:csnn-network-diagram}
\end{figure}

@@

%\subsection*{Leaky Integrate-and-Fire (LIF) Neuron Behavior}

%The Leaky Integrate-and-Fire neuron is used the in the

@@

\subsection*{Perturbation}

Another important property of each of these models is the robustness to temporal noise. We introduce an $\alpha$ parameters which represents the probability that a character is swapped for another character.

@@

\section*{Conclusion}

### 2019-05-07 paper/mybib.tex
Updates from Overleaf

>>>>

@book{sutskever2013training,
  title={Training recurrent neural networks},
  author={Sutskever, Ilya},
  year={2013},
  publisher={University of Toronto Toronto, Ontario, Canada}
}

@inproceedings{jozefowicz2015empirical,
  title={An empirical exploration of recurrent network architectures},
  author={Jozefowicz, Rafal and Zaremba, Wojciech and Sutskever, Ilya},
  booktitle={International Conference on Machine Learning},
  pages={2342--2350},
  year={2015}
}
